---
title: "Exercises for Introduction to Bayesian analysis"
author: "Benjamin Skov Kaas-Hansen"
date: "`r Sys.Date()`"
output: 
  html_document: 
    toc: yes
    highlight: haddock
    theme: paper
    # code_folding: show
editor_options: 
  markdown: 
    wrap: 72
---

------------------------------------------------------------------------

## To do for 2024 version of the course

- Part A of exercises needs to be a lot shorter
- Remove everything about simple sampling from slide deck and exercises
- Go straight to using stan instead
- Need to explain MCMC sampling in simple terms, with 1- og 2D PDF
- Consider making e.g. Shiny that produces "community-prior values" for both 
  proportion and continuous variables

------------------------------------------------------------------------

All of these exercises use the R program package. We will be needing a
few packages (all on CRAN). You can install these with the
`install.packages()` function:

```{r eval=FALSE}
install.packages(c(
  "rstan", "rstanarm", "bayesplot", "coda",
  "dplyr", "ggplot2", "purrr", 
  "adaptr"
))
```

```{r message=FALSE, class.source = 'fold-show'}
for (p in c("dplyr", "ggplot2", "purrr", "adaptr", "coda")) {
  library(p, character.only = TRUE)
}
```

The exercises revolve around two types of problems: introductory code
that you should run through to make sure you understand how to use the
function(s) and interpret the output, and additional questions to
explore and discuss to dig a bit deeper into Bayesian data analysis and
inference.

```{r echo=FALSE}
knitr::opts_chunk$set(include = TRUE, eval = TRUE)
```

## A. Bayesian analysis "by hand"

In this exercise we will take the example with the ICU patients from the
lecture and investigate what happens to the conclusions from a Bayesian
analysis when we modify the prior and have more observations. This will
also address why the straight-forward sampling approach might not be
sufficiently efficient and why specialised software would be needed.

1.  Run the code below and make sure you understand what is going on in
    each line

    ```{r include=TRUE, class.source = 'fold-show'}
    n_samples <- 10000
    n_obs <- 22 
    n_women <- 8 

    theta <- runif(n_samples) 
    cond_likelihood_draws <- rbinom(n_samples, size = n_obs, prob = theta) 
    posterior_draws <- theta[cond_likelihood_draws == n_women] 
    ```

2.  Plot a histogram of the posterior draws of theta (i.e. theta values
    you have kept because they gave rise to an outcome matching our
    original outcome)

    ```{r}
    ggplot() + 
      geom_histogram(aes(posterior_draws), bins = 10)

    ggplot() +
      stat_ecdf(aes(posterior_draws))
    ```

3.  Compute the median and 95% CrI of the posterior; provide a guess of
    the MAP

    -   Hint: R doesn't have a built-in function for finding the mode of
        a distribution

    ```{r}
    dist_mode <- function(x, ...) { # ... passed on to density()
      with(density(x, ...), x[which.max(y)])
    }

    dist_mode_bad <- function(x, n_digits = 2) { # helper function; finds mode of x
        x <- round(x, n_digits)
        ux <- unique(x)
        ux[which.max(tabulate(match(x, ux)))][1] # index in case of ties
    }

    median(posterior_draws)
    quantile(posterior_draws, c(0.025, 0.5, 0.975))
    dist_mode(posterior_draws)
    ```

4.  Now use a non-uniform distribution

    -   Hint: you are free to pick any non-uniform distribution that
        yields only values between 0 and 1; after all, it's a prior over
        a proportion (which can be neither below 0 nor above 1)

    -   Hint: a beta distribution could be interesting. Use the
        `rbeta()` function and select some interesting shape parameters
        that would express realistic prior expectations about the
        distribution between women and men
        
    ```{r}
    posterior_theta <- function(a, b, N = n_obs, z = n_women) {
        theta <- rbeta(n_samples, a, b) 
        cond_likelihood_draws <- rbinom(n_samples, size = N, prob = theta) 
        return(theta[z == cond_likelihood_draws])
    }

    prior_theta_11_11 <- rbeta(n_samples, 11, 11)
    ggplot() +
       geom_density(aes(prior_theta_11_11))
    post_theta_11_11 <- posterior_theta(11, 11)
    ```


5.  Visualise the prior distribution

    -   Hint: this can be done with e.g. histograms, density plots and
        empirical cumulative density plots

    ```{r}
    ggplot() +
      stat_density(aes(x = prior_theta_11_11), geom = "line", linetype = 2) +
      stat_density(aes(x = post_theta_11_11), geom = "line") +
      stat_density(aes(x = posterior_draws), geom = "line", colour = "red") +
      labs(x = NULL, y = NULL)
    ```

6.  Repeat the analysis from point 1 above with the new prior. Plot the
    posterior distribution, and compute the MAP and median posterior
    value.

    ```{r}
    # plot: see above
    median(post_theta_11_11)
    dist_mode(post_theta_11_11)
    ```

7.  Conclude on your analysis with words

    ```{r}
    # The distribution is pulled downward towards lower proportion of women
    ```

8.  How did the results with your own prior change compared to the
    initial analysis? Why?

9.  How many thetas are part of the posterior distribution? How does
    that compare to `n_samples`? What is the problem here?

    -   Hint: try to rerun the analysis a few times, computing each time
        the median and 95% CrI. Then, compare these summary statistics
        across runs.

    ```{r}
    length(post_theta_11_11)
    length(post_theta_11_11) / n_samples
    
    map_dfr(1:10, ~ quantile(posterior_theta(11, 11), c(0.025, 0.5, 0.975)))
    ```


10. Give a rule a thumb as to the number of effective samples (i.e. those prior 
samples giving rise to the observed outcome) you need to obtain stable results. 

    -   Hint: first define "stable results", e.g. how much the bounds of the 95% 
        CrI may change across runs.

11. Now assume that we had had 10 ICUs (all with 22 beds) but with the
    same distribution between sexes (i.e. `n_obs <- 220` and
    `n_women <- 80`). Rerun the analysis with your own prior and compare
    the results from this larger dataset to the previous analysis. How
    are they similar? How are they different?

    ```{r}
    # Data clearly dominate the prior

    posterior_draws_large <- posterior_theta(11, 11, 220, 80)
    posterior_draws_large_11_11 <- posterior_theta(110, 110, 220, 80)
    

    median(post_theta_11_11)
    median(posterior_draws_large)

    dist_mode(post_theta_11_11)
    dist_mode(posterior_draws_large)

    ggplot() +
      stat_density(aes(x = prior_theta_11_11), geom = "line", linetype = 2) +
      stat_density(aes(posterior_draws, colour = "1 ICU"), geom = "line") +
      stat_density(aes(posterior_draws_large, colour = "10 ICUs"), geom = "line")  +
      stat_density(aes(posterior_draws_large_11_11, colour = "10 ICUs, strong prior"), geom = "line") 
    ```

12. Reusing the posterior samples from exercise A6, find the percentile
    95% CrI of the posterior distribution using `quantile()`. Then,
    compute also the 95% highest posterior density interval (HPDI) and
    comment on any differences that may emerge.
    
    -   Hint: convert first to an `mcmc` object (the `coda` package has
        the functions to do this)

    -   Hint: try to plot the distribution of the posterior samples and
        draw both the percentile CrI and the HPDI.

    ```{r}
    cri <- quantile(post_theta_11_11, c(0.025, 0.975))
    hpdi <- HPDinterval(as.mcmc(post_theta_11_11), prob = 0.95)

    plot_df <- with(density(post_theta_11_11), tibble(x, dens = y)) %>% 
      mutate(
        in_95cri = between(x, cri[1], cri[2]),
        in_95hpdi = between(x, hpdi[1], hpdi[2])
      )

    ggplot(plot_df, aes(x = x, y = dens)) +
      geom_line() +
      geom_area(aes(fill = "95% CrI"), ~ filter(., in_95cri), alpha = 0.2) +
      geom_area(aes(fill = "95% HPDI"), ~ filter(., in_95hpdi), alpha = 0.2)
    
    ggplot() +
      geom_density(aes(rbeta(10000, 11, 11)))
    ```


## B. Devising and using priors

1.  Expecting an in-ICU mortality risk of 20% in patients with septic
    shock, devise a couple of priors that express this expectation with
    multiple levels of certainty. Compute the median and 95% interval

    -   Hint: you can (but don't have to!) use `find_beta_params()` from
        the `adaptr` package

    ```{r}
    map_dfr(c(0.025, 0.05, 0.1), ~ find_beta_params(0.2, ., n_dec = 2))
    ```

2.  Devise priors with the same levels of certainty but with other
    in-ICU mortality risks

    ```{r}
    expand.grid(theta = c(0.15, 0.2, 0.25), boundary = c(0.025, 0.05, 0.1)) %>% 
      bind_cols(with(., map2_dfr(theta, boundary, find_beta_params, n_dec = 2)))
    ```

3.  Assume that you are working at a production facility with quality
    control of a novel drug miraculixaban. You want to estimate the
    probability of producing a faulty batch in the coming year. In the
    previous year 365 batches were produced; none were faulty.

    Estimate the probability of faulty batch based on a frequentist
    analysis of the data. What is your estimated proportion? What is
    your forecast of faulty batches for the coming year? Can any happen?

    -   Hint: the maximum likelihood estimate of a binomial variable
        with $k$ successes (here, faulty batches) in $N$ trials (here,
        days in a year) is $\hat \theta = \tfrac{k}{N}$

    -   Hint: recall the definitions of probability from the slides

    ```{r}
    #' The maximum likelihood estimator theta (MLE) of a binomial variable with 
    #' k successes in n k/n. Because k=0, the MLE is zero, leaving no uncertainty, 
    #' and so there’s no standard error, making it impossible to actually derive 
    #' a confidence interval. So the forecast would be 0 failures.
    ```

4.  Use a Bayesian modeling approach to estimate the probability of a
    faulty batch for the coming year. How/why can the conclusion change?
    Which result is more reasonable?

    -   Hint: which prior is conjugate to a binomial likelihood (which
        is what we have here)?

    ```{r}
    #' For a beta(a, b) prior and a binomial(n, theta) likelihood, the posterior is 
    #' beta(a + k, b + n - k), because the beta distribution is a conjugate prior 
    #' to the binomial likelihood. Thus, prop_failure is the posterior 
    #' distribution of the expected probability of producing a failed drug, 
    #' assuming a beta(0.5, 0.5) prior (Jeffrey’s prior).

    prob_failure <- rbeta(10000, 0.5 + 0, 0.5 + 365)
    ggplot() + geom_density(aes(prob_failure), trim = TRUE)
    
    ggplot() + 
      geom_density(aes(rbeta(10000, 0.5 + 0, 0.5 + 365), colour = "0.5, 0.5")) +
      geom_density(aes(rbeta(10000, 1 + 0, 1 + 365), colour = "1, 1")) +
      geom_density(aes(rbeta(10000, 1 + 0, 364 + 365), colour = "1, 364"))

    # Distribution of the expected number of faults with 365 batches produced:
    ggplot() + geom_histogram(aes(prob_failure * 365), binwidth = 1)
    ```


5.  Imagine you know 10 experts with experience in the same type of
    production line as your factory. You would like to survey these
    experts and use their responses to devise an appropriate prior for
    the annual fault probability. How would you go about that?

    ```{r}
    #' One could ask them to give the number of faulty batches out of 365 
    #' produced. Then, one could compute the pooled proportion and "rescale" 
    #' to arrive at a 365-day time frame. For example:

    survey_responses <- c(1, 1, 0, 0, 2, 1, 2, 1, 1, 0)
    mean(survey_responses)

    ggplot() + 
      stat_function(fun = ~ dbeta(., 0.9, 365 - 0.9), n = 1000, xlim = c(0, 0.05))
    ```

## C. Using rstanarm

We will now perform the same analysis as we did above but using the
`rstanarm` package. We will start by loading the relevant package and
create a dataset that can be used for logistic regression.

```{r include=TRUE, message=FALSE, class.source = 'fold-show'}
library("rstanarm") 
library("bayesplot")
dta <- data.frame(woman = rep(0:1, times = c(22 - 8, 8)))
```

1.  Use the `stan_glm()` function to fit a logistic model. Remember that
    a logistic model fits parameters on the log-odds (= logit) scale.

    -   Hint: remember to set `family = binomial` to run a logistic
        regression

    -   Hint: to fit a regression model without covariates, the formula should 
        have the form `y ~ 1` where `y` is the outcome variable in the data.

```{r}
fit <- stan_glm(
  woman ~ 1, 
  data = dta, 
  family = binomial, 
  cores = 1,
  refresh = 0 # just to suppress sampler output (avoids cluttering
)
```
  

2.  Extract the posterior values and plot the posterior distribution of
    the intercept.

    -   Hint: use `as.matrix()` and `mcmc_areas()`

```{r}
posteriors <- as.matrix(fit)
as_tibble(fit)

mcmc_areas(posteriors)
```

3.  Compute the MAP and median posterior on the probability scale.

    -   Hint: recall that $g(x) = \frac{\exp(x)}{1 + \exp(x)}$ is the
        inverse function to the logistic transform
        
    -   Hint: look at `?posterior_epred`

    ```{r}
    g <- function(logit) {
      exp(logit) / (1 + exp(logit))
    }
    
    ggplot() +
      geom_density(aes(g(posteriors[, 1])))
    ```


4.  `stan_glm()` uses a zero-mean normal distribution as prior for the
    intercept by default. Thus, we might not get the same posterior
    distribution as we did above. How could we modify the prior
    distribution to make it resemble a uniform distribution? Try to
    refit the model with this uniform distribution

    -   Hint: look at `?stan_glm` and
        <https://mc-stan.org/rstanarm/articles/priors.html>

    ```{r}
    fit_uniform_prior <- stan_glm(
      woman ~ 1, 
      data = dta, 
      family = binomial, 
      prior_intercept = NULL, # uses uniform prior
      cores = 1,
      refresh = 0 # just to suppress sampler output (avoids cluttering
    )
    ```
  

5.  Try to change the prior distribution to e.g. `normal(0.0, 0.1)` and
    rerun the analysis. How does this impact the conclusion? Why?
    
    ```{r}
    fit_normal_prior <- stan_glm(
      woman ~ 1, 
      data = dta, 
      family = binomial, 
      prior_intercept = normal(0.0, 0.1),
      cores = 1,
      refresh = 0 # just to suppress sampler output (avoids cluttering)
    )
    ```

6.  Try your own prior and see how it influences the result.

    ```{r}
    #' Same as above but using whatever prior you used
    ```

7.  Prior predictive checks are like posterior predictive checks, only that they ignore the data and generates data based solely on the priors. This can often be useful to check that the priors you have specified will yield data that look reasonable for the modelling task at hand. Carry out a prior predictive check of the three models above (default prior, `normal(0.0, 0.1)` and your own). All good?

    -   Hint: see the `prior_PD` argument in `?stan_glm`.

    ```{r}
    # Create new larger dataset
    update(fit, prior_PD = TRUE)
    ```

8.  Now assume, again, that we had had data from 10 ICUs (as above).
    Construct a new data frame akin to the one above but with more
    observations. Then, rerun the analysis with your own prior and
    compare the results from this larger dataset to those of the
    previous analysis. How are they similar? How do they differ?
    
    ```{r}
    # Create new larger dataset
    dta_large <- data.frame(woman = rep(0:1, times = c(220 - 80, 80)))
    
    fit_large <- update(fit, data = dta_large) # do the same with the other fits
    summary(fit_large)
    ```
    

## D. Multiple linear regression

In this exercise we will use data from a fictive RCT with ICU patients
randomised to miraculixaban (`treatment = 1`) vs. standard of care
(`treatment = 0`) . The dataset also contains a few baseline covariates
(`age,` `woman` and `sepsis`) along with outcome variables (`icu_los`,
`hosp_los` and `dead_at_1y`).

Download the data and take a look:

```{r include=TRUE, eval=TRUE, class.source = 'fold-show'}
icu <- read.delim("https://raw.githubusercontent.com/epiben/course_adv_stats_ucph/main/bayesian/icu.tsv")
head(icu)
```

1.  Analyse the data with a regular frequentist model, estimating the
    effect of relevant covariates on the ICU length of stay. Ignore the
    `treatment` covariate for now. What are your overall conclusions?
    What are your conclusions about the impact of sepsis on ICU
    length-of-stay?

    -   Hint: relevant covariates should not contain information about
        the future

    ```{r}
    freq_fit <- lm(icu_los ~ age + sepsis + woman, data = icu)
    summary(freq_fit)
    ```

2.  Analyse the same model using a Bayesian analysis with the
    `stan_glm()` function. Try to specify reasonable priors

    ```{r cache=TRUE}
    bayes_fit <- stan_glm(
      icu_los ~ age + sepsis + woman,
      data = icu,
      seed = 4131,
      refresh = 0
    )
    
    summary(bayes_fit)
    ```

3.  Examine the fit of the model. Convergence okay?

    ```{r}
    rhat(bayes_fit)
    mcmc_dens_overlay(bayes_fit)
    mcmc_trace(bayes_fit)
    ```

4.  Use shinystan to investigate the model.

    ```{r eval=FALSE}
    launch_shinystan(bayes_fit)
    ```

5.  Increase the number of chains to, say, 8 and the double the number
    of iterations. Examine the fitted model (for example with
    shinystan). What's the impact on the results? What problems could
    more and longer chains potentially fix?

    ```{r cache=TRUE}
    bayes_fit_more_longer <- stan_glm(
      icu_los ~ age + sepsis + woman,
      data = icu,
      iter = 8000,
      chains = 8,
      seed = 4131,
      refresh = 0
    )

    summary(bayes_fit_more_longer)
    ```

6.  Look at the posterior distribution of the `sepsis` parameter. What
    is your conclusion about the effect of sepsis on ICU length of stay?
    What is the probability that the effect of sepsis is associated with 
    shortened ICU-LOS? Shortening by more than 1 day?

    ```{r}
    mcmc_areas(bayes_fit, pars = "sepsis", prob = 0.95)

    mean(as.data.frame(bayes_fit)$sepsis < 0)
    mean(as.data.frame(bayes_fit)$sepsis < -1)
    mean(as.data.frame(bayes_fit)$sepsis < -2)
    mean(as.data.frame(bayes_fit)$sepsis < -5)
    ```

7.  How might you use a survey to devise an informative prior on the
    effect of age on ICU length of stay?
    
    ```{r}
    #' One could, for example, ask respondents to distribute 20 X's on a grid with
    #' how many days longer a patient would be expected to remain admitted for
    #' each decade they are older. Then, combine all those X's into one and fit 
    #' a Normal distribution to it whose mean and sd would be the prior values.
    #' 
    #' One respondents might submit this:
    #' 
    #' ^
    #' |           X X X 
    #' |         X X X X X  
    #' |         X X X X X  
    #' |       X X X X X X X 
    #' |----------------------->
    ```
    

8.  We can use leave-one-out-cross-validation (LOO) information
    criterion (IC) to see how well different models predict the outcome.
    These LOOICs can then be compared. Like other information criteria,
    lower is better. Importantly, the LOOIC comes with a standard error.
    Fit a model without the sepsis covariate and compare the LOOIC
    between the original and the reduced models.

    -   Hint: see `?loo` and `?loo_compare`

    ```{r cache=TRUE}
    bayes_fit_wo_sepsis <- stan_glm(
      icu_los ~ age + woman,
      data = icu,
      seed = 4131,
      refresh = 0
    )

    loo_fit <- loo(bayes_fit)
    loo_fit_wo_sepsis <- loo(bayes_fit_wo_sepsis)
    loo_compare(loo_fit, loo_fit_wo_sepsis)
    ```


## E. Effect-size estimation in clinical trial

1.  A friend of yours wants to test a new treatment obelixaban against
    standard of care, in ICU patients with septic shock. Pre-trial
    evidence suggest that the 30-day mortality in this patient group
    receiving standard of care is roughly 20% (95% CrI: 5%-35%). Your
    friend's main results with respect to 30-day mortality are in the
    following 2x2 table (assume no loss to follow-up):

    <div>

    |                      | Standard of care | Miraculixaban |
    |----------------------|:----------------:|:-------------:|
    | No. alive at 30 days |       161        |      164      |
    | No. dead at 30 days  |        46        |      35       |

    </div>

    Using an appropriate prior, estimate the effect size of obelixaban
    compared to standard of care.
    
    -   Hint: recall that derived quantities, based on posterior
        parameter estimates, can be computed directly with the posterior
        samples of the parameters

    -   Hint: effect-size metrics include risk ratio (RR), risk
        difference (RD) and number-needed-to-treat (NNT)

    -   Hint: use what you learnt in exercises A6 and B1

    ```{r}
    find_beta_params(0.2, 0.05)
    post_estimates <- tibble(
      mortality_prop_control = rbeta(10000, 3 + 46, 13 + 161),
      mortality_prop_obelixaban = rbeta(10000, 3 + 35, 13 + 164),
      RR = mortality_prop_obelixaban / mortality_prop_control
    )
    quantile(post_estimates$RR, c(0.025, 0.5, 0.975))
    ```


2.  After helping your friend with their unadjusted effect-size
    estimation above, you return to your own results of the effect of
    miraculixaban against standard of care on 1-year mortality (same
    data as in section D). \
    Model the 1-year mortality risk using the four baseline covariates
    in the dataset (`treatment`, `age`, `woman` and `sepsis`) and
    default priors. What do you conclude about the conditional effects
    of miraculixaban on the 1-year mortality risk? Can you conclude
    anything about the effects of the other covariates?

    ```{r cache=TRUE}
    mortality_fit <- stan_glm(
      dead_at_1y ~ treatment + age + woman + sepsis,
      data = icu,
      family = binomial,
      seed = 4131,
      refresh = 0 # suppress fitting progress output
    )

    summary(mortality_fit, probs = c(0.025, 0.5, 0.975))

    # Convergence diagnostics
    rhat(mortality_fit)
    mcmc_dens_overlay(mortality_fit)
    mcmc_trace(mortality_fit)

    mcmc_areas(
      mortality_fit, 
      point_est = "median",
      pars = vars(-"(Intercept)"),
      prob = 0.95
    ) +
      scale_x_continuous(
        breaks = log(c(0.1, 0.2, 0.33, 0.5, 0.67, 1, 1.5, 2, 3, 5, 10)),
        labels = exp
      )
    ```

3.  Since you need to get your new drug approved by the authorities, you
    decide to use a somewhat skeptical prior on the treatment effect so
    that an estimated effect will convince even the grumpiest regulator.
    How could you devise such a skeptical prior for the effect of the
    treatment?

    -   Hint: remember that coefficients in a logistic regression are
        log-OR's and that the distribution of log-OR can be expressed in
        terms of an equivalent trial of a given size with a given event
        proportion.

    ```{r}
    r <- 0.35
    N <- 200
    (sigma <- sqrt(4/(1-r)/N + 4/r/N))

    ggplot() + 
      geom_density(aes(rlnorm(1000, meanlog = 0, sdlog = sigma))) +
      scale_x_log10(breaks = c(0.33, 0.5, 0.67, 0.8, 1, 1.25, 1.5, 2, 3))
    ```

4.  Rerun the logistic regression with this new informative prior.
    What's the impact on the (estimated) conditional treatment effect?

    ```{r message=FALSE, cache=TRUE}
    # Only difference is the scale of the first parameter
    priors <- normal(
      location = c(0, 0, 0, 0),
      scale = c(sigma, 2.5, 2.5, 2.5)
    )

    mortality_fit_skeptical <- stan_glm(
      dead_at_1y ~ treatment + age + woman + sepsis,
      data = icu,
      prior = priors,
      family = binomial,
      seed = 4131,
      refresh = 0 # suppress fitting progress output
    )

    summary(mortality_fit_skeptical, probs = c(0.025, 0.5, 0.975))

    prior_summary(mortality_fit_skeptical)

    map(list(mortality_fit, mortality_fit_skeptical), as.data.frame) %>% 
      map_dfc("treatment") %>% 
      setNames(c("default", "skeptical")) %>% 
      ggplot() +
        geom_density(aes(x = exp(default), colour = "default")) +
        geom_density(aes(x = exp(skeptical), colour = "skeptical")) +
        scale_x_log10(breaks = c(0.1, 0.2, 0.33, 0.5, 0.67, 1, 1.5, 2, 3, 5, 10)) +
        labs(x = "OR", y = NULL)
    ```

5.  Compute average treatment effect using G-computation

    -   Hint: predict the posterior probabilities of death for all patients if 
        they had all received the miraculixaban and if they had all received 
        standard of care Take a look at ?posterior_epred
    -   Hint: here are two links for good introductions to G-computation:
        https://github.com/kathoffman/causal-inference-visual-guides/blob/master/visual-guides/G-Computation.pdf, 
        https://doi.org/10.1093%2Fije%2Fdyw323

    ```{r}
    posterior_marginal_estimates <- map_dfc(
      c(miraculixaban = 1, control = 0),
      ~ rowMeans(posterior_epred(mortality_fit, mutate(icu, treatment = .)))
    ) %>% 
      mutate(
        risk_ratio = miraculixaban / control,
        risk_diff = control - miraculixaban,
        NNT = 1/risk_diff
      )

    ggplot(posterior_marginal_estimates) + 
      geom_density(aes(risk_ratio)) 
    ```


## F. Coding up linear regression in Stan

**NB! We didn't make it here**

Build and run your own Stan model of a linear regression as the one in
D2 above. Analyze the full data with the lung capacity as a Stan model
and run it. Argue for all the choices that you have made so far
regarding priors, likelihood, etc. Recall from the lecture that a basic
Stan program contains three sections: data, parameters, and model. Below
is the sample code for a standard linear regression analysis:

``` stan
data {
  int<lower=1> N;
  vector[N] y;
  vector[N] x;
}

parameters {
  real intercept;
  real beta;
  real<lower=0> sigma; // note lower limit
}

model {
  // Prior distributions
  intercept ~ normal(0, 10);
  beta ~ normal(0, 10); 
  sigma ~ cauchy(0, 2.5);
  
  // Likelihood
  y ~ normal(intercept + beta * x, sigma);
}
```

1.  Modify this code to use all covariates and run the analysis

    -   Hint: start by including and using the two numeric variables
        first.

    -   Hint: then include the categorical variables as numeric
        variables (with values either 0 or 1). This works because the
        categorical predictors are binary.

## G. Resources

- Book - Richard McElreath: Statistical Rethinking [https://xcelab.net/rm/statistical-rethinking/](https://xcelab.net/rm/statistical-rethinking/). Excellent introduction to Bayesian analysis. You really should start here.
- Article - Michael Betancourt: A Conceptual Introduction to Hamiltonian Monte Carlo [https://arxiv.org/pdf/1701.02434.pdf](https://arxiv.org/pdf/1701.02434.pdf). Excellent paper on HMC; it is a bit technical, but even if you skip the most technical parts and equations, there's lots of learning in there.
- Article - Goligher et al.: Extracorporeal membrane oxygenation for severe acute respiratory distress syndrome and posterior probability of mortality benefit in a post hoc bayesian analysis of a randomized clinical trial [https://doi.org/10.1001/jama.2018.14276](10.1001/jama.2018.14276)
- Book - James V. Stone: Bayes' Rule: a tutorial introduction to Bayesian analysis. A bit old but with a fun-ish take on things and not too much math.
- Book - Andrew Gelman et al.: Bayesian Dataz Analysis. The Bible but really not accessible and in no way the place to start for most people.